---
title: "p8105_hw2_sal2222"
author: "Stephen Lewandowski"
date: "October 4, 2018"
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
```

# Problem 1


First, I will read and clean the NYC Transit data.

```{r}
nyc_subway_data <- read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, 
         entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE)) 
 
nyc_subway_data
  
```
Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance?
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? How many are ADA compliant?




# Problem 2


# Problem 3